---
title: "pre_dbem_analysis"
author: "Juliano Palacios"
date: "25/07/2023"
output: html_document
---

```{r}
library(MyFunctions)

my_lib(
  c(
    "sf",
    "ncdf4",
    "metR",
    "tidyverse",
    "doParallel",
    "foreach"
  )
)

```


# COBALT2 data

From Matthias B.

the 0.5° variant is now available as Secondary Input Data at DKRZ under

`/work/bb0820/ISIMIP/ISIMIP3a/SecondaryInputData/climate/ocean/[obs|ctrl]clim/global/monthly/historical/GFDL-MOM6-COBALT2/`


## Data needed

Translate ISIMIP data to DBEM:

### get data from server

```{r}

# data address
# /work/bb0820/ISIMIP/ISIMIP3a/SecondaryInputData/climate/ocean/ctrlclim/global/monthly/historical/GFDL-MOM6-COBALT2

# Call data
# scp -r b381132@levante.dkrz.de:

# Variables needed
gfdl-mom6-cobalt2_ctrlclim_o2-bot_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_o2-surf_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_ph-bot_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_ph-surf_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_so-bot_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_so-surf_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_tob_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_tos_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_siconc_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_uo_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_vo_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_intppdiat_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_intppdiaz_30arcmin_global_monthly_1961_2010.nc
gfdl-mom6-cobalt2_ctrlclim_intpppico_30arcmin_global_monthly_1961_2010.nc

```



```{bash}


# Shell loop

#!/bin/bash

# Replace 'your_remote_username' and 'your_remote_server_ip' with the actual username and IP address of the remote server.
# Replace '/path/to/files/' with the actual path to the directory where the files are located on the remote server.
# Replace '/path/on/your/computer/' with the local directory where you want to save the transferred files.

files=( "gfdl-mom6-cobalt2_obsclim_ph-bot_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_o2-bot_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_o2-surf_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_ph-bot_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_ph-surf_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_so-bot_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_so-surf_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_tob_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_tos_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_siconc_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_uo_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_vo_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_intppdiat_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_intppdiaz_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_obsclim_intpppico_30arcmin_global_monthly_1961_2010.nc")


# For ctrlclim data
#files=("gfdl-mom6-cobalt2_ctrlclim_o2-bot_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_o2-surf_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_ph-bot_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_o2-bot_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_o2-surf_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_ph-bot_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_ph-surf_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_so-bot_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_so-surf_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_tob_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_tos_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_siconc_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_uo_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_vo_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_intppdiat_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_intppdiaz_30arcmin_global_monthly_1961_2010.nc" "gfdl-mom6-cobalt2_ctrlclim_intpppico_30arcmin_global_monthly_1961_2010.nc")




# For fishing effort
files=("effort_histsoc_1841_2010.csv"
"gridded_industrial_effort_histsoc_1961_2010.csv"
"gridded_artisanal_effort_histsoc_1961_2010.csv")

scp -r b381132@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3a/InputData/socioeconomic/fishing/histsoc/"$file" ./

for file in "${files[@]}"
do
    scp -r b381132@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3a/SecondaryInputData/climate/ocean/obsclim/global/monthly/historical/GFDL-MOM6-COBALT2/"$file" ./
    
done


scp -r b381132@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3a/InputData/climate/ocean/obsclim/global/monthly/historical/GFDL-MOM6-COBALT2/"$file" ./

#Save the script to a file and make it executable with the following command:

#chmod +x transfer_files.sh

#Then, you can run the script using:

#./transfer_files.sh


 #your_remote_username@your_remote_server_ip

#ssh-copy-id b381132@levante.dkrz.de

```


### Transform data to DBEM like

```{r}

# Set path for data
data_path <- my_path("G","Climate/cimip6/COBALT2/obsclim")

# List files

climate_data <- list.files(data_path,full.names = T)

```

## Get DBEM grid on data


```{r}

grid_dbem <- function(shapefile, dbem_grid, out = "df"){

  dbem_grid <- sf::st_as_sf(dbem_grid,
                            coords = c("lon","lat"),
                            crs = 4326)

  # if supplied a grid will rasterize
  if(inherits(shapefile, "sf")== FALSE){

    shapefile <- sf::st_as_sf(shapefile,
                              coords = c("lon","lat"),
                              crs = 4326) 
  }else{

    # Make sure shapefile is in the same coordinate system
    shapefile <- shapefile %>%
      st_set_crs(4326)
  }

  # Merge both
  if(out == "sf"){
    merge_grid <- sf::st_join(dbem_grid,
                              shapefile,
                              join = st_intersects)
  }else{
    merge_grid <- sf::st_join(dbem_grid,
                              shapefile,
                              join = st_intersects) %>%
      as.data.frame() %>%
      select(-geometry)
  }
  return(merge_grid)
}



```


```{r}
# Use one dataset to grid

# Load nc data

shapefile <- ReadNetCDF(climate_data[1]) %>%
  filter(time == 720) %>%
  tibble::rowid_to_column() %>% 
  select(rowid,lon,lat)


```

```{r}

variable_list <- data.frame(
  isimip_name = c(
    "intppdiat",
    "intppdiaz",
    "intpppico",
    "o2-bot", 
    "ph-bot",
    "so-bot",
    "tob",
    "siconc",
    "o2-surf",
    "ph-surf",
    "so-surf",
    "tos",
    "vo"
  ),
  dbem_name = c(
    "diat_",
    "diaz_",
    "pico_",
    "O2_btm_",
    "htotal_btm_",
    "Salinity_btm_",
    "bot_temp_",
    "IceExt_",
    "O2_surf_",
    "htotal_surf_",
    "Salinity_surf_",
    "SST_",
    "AdvectionV_")
)



to_dbem <- function(data,yr,var,forcing){
  
  x <- data %>% 
    filter(year == yr) %>%
    ungroup() %>% 
    select(variable)
  
  
  name <- paste0("~/Library/CloudStorage/OneDrive-UBC/Data/Climate/cimip6/COBALT2/to_dbem/",forcing,"/",var$dbem_name,yr,".txt")
  write.table(x, file=name, sep="\t", col.names = F, row.names = F)
 gc() 
}



```



## Routine


```{r}


# for(i in 1:13){
for(i in c(7,10,12)){
  # i = 5
  print(climate_data[i])
  
  suppressWarnings(
    data <- ReadNetCDF(climate_data[i]) %>% 
      rename_with(~ "variable", 5) %>% # change to 5 for i = 5,7,10,12 due to data inconsistency
      left_join(shapefile,
                by = c("lon","lat")) %>% 
      mutate(date = ymd_hms("1901-01-01 00:00:00") + months(time),
             year = year(date)) %>% 
      group_by(rowid,year,lon,lat) %>% 
      summarise_at(c("variable"),
                   mean,
                   na.rm = T) %>% 
      mutate(variable = ifelse(is.na(variable), -9999, variable))
  )
  
  
  var <- variable_list %>% 
    filter(str_detect(climate_data[i],isimip_name))
  
  mclapply(seq(1961,2010,1), to_dbem, data = data, var = var, forcing = "obsclim")
  
}



```

## test outputs

They all look good

```{r}

dbem_coords <- my_data("dbem_coords")

files <- list.files("~/Library/CloudStorage/OneDrive-UBC/Data/Climate/cimip6/COBALT2/to_dbem/ctrlclim",pattern = "SST",full.names = T)

suppressMessages(
  clim_data <- bind_cols(
    lapply(files, read_csv, col_names = F)
  )
)

colnames(clim_data) <- seq(1961,2010,1)


gridExtra::grid.arrange(
  clim_data %>% 
    bind_cols(dbem_coords) %>% 
    select(lat,lon,index,`2000`) %>% 
    filter(`2000` > -100) %>% 
    # View()
    ggplot() +
    geom_tile(
      aes(
        x = lon,
        y = lat,
        fill = `2000`
      )
    ),
  clim_data %>% 
    bind_cols(dbem_coords) %>% 
    gather("year","value",`1961`:`2010`) %>% 
    filter(value > -100) %>% 
    group_by(year) %>% 
    summarise(mean_v = mean(value,na.rm=T)
    ) %>% 
    ggplot() +
    geom_line(
      aes(
        x = as.numeric(year),
        y = mean_v
      )
    )
)

```

## Fix -9999

Forgot to transform NAs to -9999

```{r}


dbem_files <- list.files("~/Library/CloudStorage/OneDrive-UBC/Data/Climate/cimip6/COBALT2/to_dbem/ctrlclim",full.names = T)


for(i in 200:length(dbem_files)){
  
  correct <- read.table(dbem_files[i]) %>% 
    mutate(V1 = ifelse(is.na(V1), -9999, V1))
  
  write.table(correct, file=dbem_files[i], sep="\t", col.names = F, row.names = F)
}

```

## Manually extract u data cus is too big 

```{r}


forcing <- "obsclim"

file_path <- list.files(paste0("~/Library/CloudStorage/OneDrive-UBC/Data/Climate/cimip6/COBALT2/",forcing,pattern = "uo",full.names = T))

nc <- nc_open(file_path)

# Extract relevant information
lon <- ncvar_get(nc, "lon")
lat <- ncvar_get(nc, "lat")
time <- ncvar_get(nc, "time")
uo <- ncvar_get(nc, "uo", start = c(1, 1, 1, 1), count = c(-1, -1, 1, -1))  # Extract only the first Z layer

# Close the NetCDF file
nc_close(nc)

# Get the time in years
years <- as.integer(time / 12) + 1901

# Create a data frame
df <- expand.grid(
  lon = lon,
  lat = lat,
  years = years
) %>% 
  left_join(shapefile,
              by = c("lon","lat")) %>% 
  left_join(dbem_coords)

# Add the Sea Water X Velocity data
df$Velocity <- as.vector(uo)

# Print the resulting data frame
print(df)


# Visualize data
df %>%
  filter(years == 2000) %>%
  group_by(lon,lat) %>%
  summarise(mean = mean(Velocity, na.rm = T)) %>%
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = mean
    )
  )


# Save data for DBEM (Remeber to convert NAs to -9999)

ou_data <- df %>%
  group_by(year = years,index) %>%
  summarise(mean = mean(Velocity, na.rm = T)) %>% 
  mutate(variable = ifelse(is.na(mean), -9999, mean))


to_dbem <- function(data,yr,var,forcing){
  
  x <- data %>% 
    filter(year == yr) %>%
    ungroup() %>% 
    select(variable)
  
  
  name <- paste0("~/Library/CloudStorage/OneDrive-UBC/Data/Climate/cimip6/COBALT2/to_dbem/",forcing,"/",var,yr,".txt")
  write.table(x, file=name, sep="\t", col.names = F, row.names = F)
 gc() 
}


mclapply(seq(1961,2010,1), to_dbem, data = ou_data, var = "AdvectionU_", forcing = forcing)

```

## Transform totalphy


totalphy2= (intppdiaz+intppsmall+intppico)*0.1 + diatoms

```{r}

# Get data paths
diat <- list.files("~/Library/CloudStorage/OneDrive-UBC/Data/Climate/cimip6/COBALT2/to_dbem/ctrlclim", pattern = "diat", full.names = T)
diaz <- list.files("~/Library/CloudStorage/OneDrive-UBC/Data/Climate/cimip6/COBALT2/to_dbem/ctrlclim", pattern = "diaz", full.names = T)
pico <- list.files("~/Library/CloudStorage/OneDrive-UBC/Data/Climate/cimip6/COBALT2/to_dbem/ctrlclim", pattern = "pico", full.names = T)

for(y in 1:length(diat)){
  
  # Read data
  npp_df <- bind_cols(
  read.table(diat[y]),
  read.table(diaz[y]),
  read.table(pico[y])
  )
    colnames(npp_df) <- c("diat","diaz","pico")
    
  
  totalphy_df <- npp_df %>% 
    mutate(totalphy = (diaz+pico)*0.1+diaz) %>% 
    mutate(totalphy = ifelse(is.na(totalphy),-9999,totalphy)) %>% 
    select(totalphy)
  
  name <- paste0("~/Library/CloudStorage/OneDrive-UBC/Data/Climate/cimip6/COBALT2/to_dbem/ctrlclim/","totalphy2",str_sub(diat[y],99,102),".txt")
  write.table(totalphy_df, file=name, sep="\t", col.names = F, row.names = F)
  
}


read.table("~/Library/CloudStorage/OneDrive-UBC/Data/Climate/cimip6/COBALT2/to_dbem/ctrlclim/totalphy22000.txt") %>% 
  bind_cols(dbem_coords) %>% 
  filter(V1 > -100) %>% 
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = V1
    )
  )

```

# Old code

### Testing COBALT and GFDL85

U is giving me a lot of trouble, how different is V? Can I just substitute it? No, you can't. They are very different.



```{r}

gfdl_path <- list.files(my_path("G",extra_path = "Climate/cimip6/C6GFDL85/"),pattern = "U", full.names = T)

gfdl_data <- bind_cols(
  lapply(gfdl_path, read_csv)
)
colnames(gfdl_data) <- seq(1851,2100,1)

gfdl_data %>% 
  select(`2000`) %>% 
  filter(`2000` > -100) %>% 
  rowid_to_column("index") %>%
  left_join(my_data(data = "dbem_coords")) %>% 
  # select(lat,lon,`1967`) %>% 
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = `2000`
    )
  )

gfdl_clean <- gfdl_data %>% 
  gather(year,value,`1961`:`2010`) %>% 
  select(year,value)


cobalt_path <- list.files(my_path("G","Climate/cimip6/COBALT2/to_dbem/ctrlclim/"), pattern = "V", full.names = T)

cobalt_data <- bind_cols(
  lapply(cobalt_path, read_csv)
)
colnames(cobalt_data) <- seq(1961,2010,1)

cobalt_clean <- cobalt_data %>% 
  gather(year,value,`1961`:`2010`) %>% 
  select(year,value)

cobalt_data %>% 
  rowid_to_column("index") %>% 
  left_join(my_data(data = "dbem_coords")) %>% 
  select(lat,lon,`1967`) %>% 
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = `1967`
    )
  )


# Compare datasets

gfdl_clean %>% 
  mutate(model = "gfdl") %>% 
  bind_rows(cobalt_clean %>% mutate(model = "cobalt")) %>% 
  filter(value> -9000) %>% 
  group_by(year,model) %>% 
  summarise(mean = mean(value,na.rm =T),
            max = max(value,na.rm =T),
            min = min(value,na.rm =T)
            ) %>% 
  gather("var","value",mean:min) %>% 
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = value,
      color = model
    )
  ) +
  facet_wrap(~var,
             scales = "free")


```
```{r}



```

